---
title: "uniprotREST"
output: github_document
---

<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

This package provides an R wrapper for the 
[UniProt website REST API](https://www.uniprot.org/help/api).

## Installation

Install the latest version from R-universe:

```{r, eval=FALSE}
install.packages("uniprotREST", repos = "https://csdaw.r-universe.dev")
```

Or install the latest development version from GitHub:

```{r, eval=FALSE}
remotes::install_github("csdaw/uniprotREST")
```

## Documentation

Read the full documentation [here](https://csdaw.github.io/uniprotREST/articles/uniprotREST.html).

## Quick start

```{r}
library(uniprotREST)
```

### ID mapping with `uniprot_map()`

Map to/from UniProt IDs. This function wraps the 
[ID mapping](https://www.uniprot.org/help/id_mapping) API endpoint.

```{r}
# Proteins of interest (from 3 different taxa)
ids <- c("P99999", "P12345", "P23456")

# Get accessions, gene names and sequence lengths
result <- uniprot_map(
  ids = ids, 
  from = "UniProtKB_AC-ID",
  to = "UniProtKB",
  format = "tsv",
  fields = c("accession", "gene_primary", "length")
)

result
```

### Querying UniProt with `uniprot_search()`

Perform text searches against UniProt databases.
This function wraps the [Query](https://www.uniprot.org/help/api_queries) 
API endpoint.

```{r}
# Get human glycoproteins less than 100 amino acids long

result <- uniprot_search(
  query = "(proteome:UP000005640) AND (keyword:KW-0325) AND (length<100)",
  database = "uniprotkb",
  format = "tsv",
  fields = c("accession", "gene_primary")
)

head(result)
```

### Retrieving an entry with `uniprot_single()`

Download the full entry for a single protein. This function wraps the 
[Retrieve](https://www.uniprot.org/help/api_retrieve_entries) API endpoint.

```{r}
# Human cytochrome C
result <- uniprot_single(
  id = "P99999",
  database = "uniprotkb",
  format = "json",
  verbosity = 0
)

str(result, max.level = 1)
```

## Metadata

- Please [report any issues or bugs](https://github.com/csdaw/uniprotREST/issues)
  ideally with a [reproducible example](https://reprex.tidyverse.org/)
- This project is released with a [Contributor Code of Conduct](https://csdaw.github.io/uniprotREST/code_of_conduct.html). 
  By participating in this project you agree to abide by its terms.
- Author: Charlotte Dawson
- License: MIT
